### 2019-08-24 abstract.tex
first commit

This system was \#1 in the world for a very long time, and has remained
a critically important computer system. It defied scale with 18,688
individual compute nodes, and delivered tens of billions of computing
hours to DOE mission-critical programs for nearly 7 years.