### 2019-08-24 acknowledgments.tex

\section{Acknowledgments}
\label{sec:ack}
\fix{Standard and all}