### 2020-04-20 discussion.tex
Worked mostly on the discussion section.

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{section:discussion}

\fix{A short summary of the observations in this paper to facilitate the discussion below.}

The experienced unexpected reliability issues with the Titan supercomputer raise
several questions and discussion points. First, there is the mitigation component
that keeps the system operating at an acceptable level and possibly even restores
some, if not all, of the lost capability. Second, there is the long-term planning
component that addresses such scenarios in future-generation supercomputers.

In Titan, the mitigation component included replacing about 11,000 GPUs and
changing the job scheduling strategy~\cite{8665764}. Replacing about 59\% of
Titan's 18,688 GPUs helped to improve productivity by restoring some of the lost
capability, but it was a costly and time-consuming effort. Changing the job
scheduling strategy to run larger jobs on more reliable and smaller jobs on less
reliable nodes played a crucial role in maintaining productivity at reduced
capability.

The options of replacing failed components and employing reliability-aware
resource management may not be available or cost effective for other
supercomputing centers dealing with similar unexpected reliability issues in
their systems. Replacement components may not be readily available and may have
to be manufactured, which can be impossible for technology that is no longer
supported by the original manufacturer. There is also a financial aspect, such
as service contract and/or warranty expiration and the operational budget of a
supercomputing center not having significant funds for replacement costs. There
is also the time and financial cost for root-cause analysis and developing a
hardware and software mitigation strategy in the first place. Reliability-aware
resource scheduling also has its limitations, as the network architecture needs
to be taken into account. Titan's 3D torus network creates certainly more
challenges for efficient job allocations, than the fat tree network of Titan's
successor, Summit~\cite{olcf:summit}. Nodes associated with the same job need to
be close to each other on the network for performance. Alternatively, partially
or completely replacing the aging and failing supercomputer with a new system,
even if this solution offers less capability, may be more cost effective in the
end.

Other mitigation components for Titan included matching the checkpointing
interval of applications with the system's current
reliability~\cite{bautista-gomez16reducing, 6903564}. The experienced issue here
was the lack of automated communication of system reliability information and
the lack of flexibility in application-level checkpoint/restart implementations.
Constantly measuring and reporting current system reliability is not an automated
real-time task in currently deployed supercomputers. Applications are typically
not designed to skip a checkpoint when system reliability is good or take more
frequent checkpoints when system reliability is getting worse.

The long-term planning component is one of the biggest lessons learned from the
Titan reliability experience. Today's supercomputers are designed to deal with
expected reliability issues. However, history has shown~\cite{geist12kill} that
unexpected reliability issues do occur and do have a significant impact. Vendors
and manufacturers obviously can not mitigate against all possible reliability
threats, however, a resilience strategy is needed for future-generation systems
that is able to deal with emerging unexpected reliability threats in a
reasonable and cost effective way. In addition to mitigation, better support for
automated real-time reliability monitoring and reporting is needed, including
for root-cause analysis.

\fix{\begin{itemize}
\item Cray and NVIDIA take care of the hardware and some systems and driver
  software so moving and troubleshooting hardware is their cost.
\item Checkpoint strategies, allocation strategies, job resource
  management, are all under OLCF control. What can OLCF do/change to
  mitigate fallout in terms of delivering computing to customers when
  facing machine disasters like the two episodes (2013 and 2016) on Titan?
\end{itemize}}

\fix{RA: Recommendation for system operators if they see similar 
wide-spread failures in their system well before the expected lifetime 
of the machine. Should they replace all components or just partially
upgrade the machine. Tradeoff?}