### 2020-05-27 acknowledgments.tex
finished reviewer 3

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
\label{section:acknowledgments}

We are extremely grateful to four anonymous reviewers who provided detailed and
insightful comments on our initial submission, which resulted in a
greatly improved presentation and deeper conclusions.

This work was supported by \doubleblind{the U.S. Department of Energy, Office of Science,
Office of Advanced Scientific Computing Research, program managers Robinson Pino
and Lucy Nowell, under contract number DE-AC05-00OR22725.}{doubleblind}