### 2020-06-04 reviews/response.txt
done today

We thank all the reviewers for their exceptionally helpful and
thorough comments. Thorough reviews take time and interest! We have
incorporated suggested changes in our revised version and marked the
revised text in blue. Here, we respond to reviewers' "Comments for
Revision" and due to limited space explain mostly only those detailed
comments that are not marked or incorporated in our revised paper
version.

Overall, there were two substantive additions in addition to numerous
smaller cosmetic and presentation improvements. The first is a
corroboration of our airflow implications on temperature with data
from another publication. The second is finding a contributing factor
to temperature in job scheduling