### 2020-06-05 reviews/response.txt
corrosion sentence in intro

We thank the reviewers for their exceptionally helpful and thorough
comments. Thorough reviews take time, interest, and expertise! Here,
we respond to reviewers' "Comments for Revision", marked by (C), and
due to limited space explain only a selection of detailed comments,
marked by (D). However, we have addressed all of the reviewers
comments in our revised paper either by making the requested change or
by providing additional explanation.

Our revision includes two substantive additions and numerous
improvements to explanations, graphics, and presentation. The first
substantive addition is a corroboration of our airflow implications on
temperature with temperature data from another publication. The second
substantive addition is a new result that job scheduling and folded
torus interconnect were contributing factors to temperature patterns,
and explain a peculiar interleaving column hazard variability.

Reviewer 1:

(C) Figure 2 is improved and corrected. The inventory dates are the same
for old and new batches, except that new units begin to appear in 2016 and
this is evident from the year colors.

Figure 7 is now split between old and new batches.

We now have location-specific information for average temperature
across cages and it agrees with our airflow effect on temperature
interpretation (V.-6th paragraph).

(D)

Reviewer 2:

(C) We have strengthened our explanations