### 2020-06-05 reviews/response.txt
some rephrasing

We thank the reviewers for their exceptionally helpful and thorough
comments. Thorough reviews take time, interest, and expertise! Here,
we respond to reviewers' "Comments for Revision", marked by (C), and
due to limited space explain only a selection of detailed comments,
marked by (D). However, we have addressed all of the reviewers
comments in our revised paper either by making the requested change or
by providing additional explanation.

Our revision includes two substantive additions and numerous
improvements to explanations, graphics, and presentation. The first
substantive addition is a corroboration of our airflow implications on
temperature with temperature data from another publication. The second
substantive addition is a new result that job scheduling and folded
torus interconnect were contributing factors to temperature patterns,
and explain a peculiar interleaving column hazard variability.

We have added more explantion to most figures, making them more
self-contained and relevant to our conclusions.

Reviewer 1:

(C)
Figure 2 is improved and corrected. The inventory dates are the same
for old and new batches, except that new units begin to appear in 2016 and
this is evident from the year colors.

Figure 7 is now split between old and new batches.

We now have location-specific information for average temperature
across cages and it agrees with our airflow effect on temperature
interpretation (V.-6th paragraph).

(D)

Reviewer 2:

(C)
We have strengthened our explanations in several sections and
particularly around our airflow explanation in Fig.4. Both substantive
improvements described above relate to this point, on one hand
strengthening it with temperature data, and on the other hand also finding
a more nuanced impact of scheduling.

(D)
Section 3: see changes in first paragraph.

Reviewer 3:

(C)
Generalization: This analysis is already generating interest from
Summit operations. A lesson learned is that Titan was not well
instrumented for data collection - thus our difficulty in obtaining
temperature data. Frontier will likely have even more
instrumentation.

Novelty in graphics: Figures 1 and 5 were specifically designed for
this data using R's amazingly high-level and flexible ggplot2
concepts. Figures 10-13 are standard in R's survival analysis
packages.

(D)

Reviewer 4:

(C)

(D)