### 2020-06-05 intro.tex
novelty notes

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
The Cray XK7 Titan supercomputer~\cite{olcf:titan} was the \#1 system
in the world for a very long time~\cite{top500list}, and has remained
a critically important computer system through the end of its life in
the Summer of 2019. It defied scale with 18,688 individual compute
nodes and delivered tens of billions of computing hours to the
U.S. Department of Energy mission-critical programs for nearly 7
years.

From a realiability perspective, Titan was a very interesting
machine. Its operation was forced to execute three very significant
rework cycles, two on the mechanical assembly affecting the PCIe
connector from the GPU daughtercard to the motherboard, and one to
replace about 11,000 GPU assemblies because of a failing resistor on
their printed circuit board. We write primarily about the GPU
operation epoch that includes this last rework cycle. This epoch of
nearly 6 years includes Titan's most stable and failure free period
and contains the most reliable data on GPU operation.

\begin{figure*}[tb]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{\rev{Volume of GPU swaps detected on Titan at individual
    inventories (narrow blue) and yealy sum totals for 2014 and later
    (wide gray) over its entire lifetime. High swap volumes prior to
    2014 reflect two major rework cycles on mechanical assembly issues
    at PCIe connectors that can be characterized as a break-in
    period. Our analysis in this paper excludes the break-in period
    and is focused on units in place by the end of 2013 or swapped in
    during the years that follow}.}
  \label{fig:chronology}
\end{figure*}
Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories. The first two rework cycles before 2014
involving blade mechanical assemblies can be characterized as a
break-in period on a new system that is first of its kind and pushes
many technological boundaries. \rev{This early period included
  extensive field engineering and experimentation, down times, as well
  as temporal and completeness gaps in inventory runs due to down
  times. We include the early swap data in this figure only for
  completeness and to underscore the massive amount of hardware work
  required to field a world's largest supercomputer. However, as we
  focus on GPU operation after the second rework cycle we exclude this
  early period from further analysis.} Late 2013 begins a very long
period of high reliability and stable operation with very few issues
until about mid 2016, when GPU failures begin to rise. This results in
the final rework cycle, replacing 11,000 GPUs throughout much of 2017.

Finding the root cause for the 2016 failures took a great deal of
testing and the involvement of materials science and microscopy
researchers. The failures were traced to a resistor on the GPU circuit
board (not the GPU chip iself) due to silver sulphide
corrosion. \fix{Was anything published about this?  If not, let's use:
  Growth of such corrosion products in ambient air on microelectroncs
  parts is described in \cite{Vargas2009TheCO} and is cosistent with
  failures starting only after a critical amount of corrosion builds
  up, a situation that matches the experience on Titan}.

Throughout this paper, when we refer to replacing the GPU, we refer to
the entire circuitboard along with its GPU chip. Our analysis focuses
on GPU boards that were installed in the second rework cycle and
later, essentially units installed near the beginning of 2014 and
later. Figure~\ref{fig:chronology} is the only one involving data
prior to 2014. All analyses and figures that follow use data from 2014
onward. We make the data as well as our analysis R codes and Python
codes to reproduce the graphics in this paper publicly available (see,
\cite[repository to be added in final
version]{repositoryreference,rcodes,pycodes}).  While the data is only
a 7 MB file, due to the size of Titan it represents over 100,000
collective years of GPU operation and may be the largest publicly
available GPU reliability data set. Our hope is that others will use
it for reliability and survival analysis research and go beyond the
results presented here.

We begin by describing related work in Sec.~\ref{section:related}. Our
data collection is described in Sec.~\ref{section:dataprep} and the
data cleaning, checking, and lifetime summarization process in
Sec.~\ref{section:dataclean}. \rev{Here we develop novel
  visualizations of GPU life times and errors that were needed to
  understand the data and our data processing decisions.} A time
between failures (TBF) analysis is given in
Sec.~\ref{section:tbf}. Survival analysis \rev{(SA)} based on
Kaplan-Meier modeling (KM) and Cox regression modeling (CR) of
relative hazard rates in various locations are in
Sec.~\ref{section:survival}. The goals of a TBF analysis and of SA are
different: TBF studies largely \rev{show} GPU reliability affects the
machine, whereas SA studies largely how the machine affects GPU
reliability. \rev{The application of SA methods is novel in the HPC
  reliability context. It has roots in biostatistics and epidemiology,
  where it is used to discover causes of disease.} Finally, we discuss
our main conclusions and recommendations in
Sec.~\ref{section:discussion}.